Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-10-17 TACAS2018/paper.tex
Fixes

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable:
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
%
where $t \in \mathbb R_0^+$, where $A$ and $B$ are matrices that 
specify the continuous plant, and with initial states set as $x(0)$.

@@

In this work, equations~\eqref{eq:ode1} and~\eqref{eq:ode2} is soundly discretized in time (see Appendix~\ref{sec:cont_aa}) \todo{or use another paper as reference and remove sec .1?} into
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k\\
y_{k} = C_d x_k + D_d u_k
\end{align} 
%
where $k \in \mathbb N$ and $x_{0}=x(0)$ is the initial state. 
$A_d$, $B_d$, $C_d$ and $D_d$ denote the matrices that describe the discretized plant dynamics, whereas $A$, $B$, $C$ and $D$ denote the continuous plant dynamics.  
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
\todo{$\mathbb{R}^p$?}
The traces of the model starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs restricted to $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big( \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big)\;. 
\end{equation}
%
\todo{$\mat{G}, \vec{h}$?}
%
We extend the notation above to convex sets of initial conditions and inputs ($X_0$ and $U$), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x}_0 \in
X_0 \wedge \vec{u} \in U \wedge \vec{x} = \tau(\vec{x}_0,\vec{u})\}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore write $\tau^n(X_0,U)$ to denote the set of states reached from
$X_0$ via $\tau$ in $n$ steps (\emph{$n$-reach set}), i.e.\ for $n\geq 1$
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid &\vec{x}_0 \in X_0 \wedge \forall k\in [0,n-1]\nonumber\\ 
 &\vec{u}_{k} \in U \wedge \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and
vector sums preserve convexity, the sets $X_n = \tau^n(X_0,U)$ are also
convex.

@@

 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, we have
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \text{where }\mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{and } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  We call $\vec{x}_T$ a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 %% We can apply Th.~\ref{eq:continuous_tube_dyn} to discretized systems without inputs in order to
 %% obtain a witness ..
%
\todo{Can the corallary be motivated? The general motivation of the section was observing the system at times that are not necessarily integer multiple of the sampling time...} 
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0 : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in observing systems with parametric
 inputs, described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$.
 For such systems, a witness $\vec{x}_T$ is computed next.
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
\end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in (0,\infty)$ is a witness of the system at time $kT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

@@

\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T, \mat{B}_T$ as per \eqref{eq:continuous_tube_dyn} and $\mat{A}_d, \mat{B}_d$ describing the discretized plant dynamics is a witness of the system
 \begin{align}
 \dot{\vec{x}}(t) &= \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\nonumber\\
 \vec{u}_t&=\mat{K}\vec{x}_k  \mid  kT_s \leq t < (k+1)T_s
 \end{align}
 \end{theorem}
%
\todo{\eqref{eq:continuous_tube_dyn} does not define $\mat{B}_T$}
\todo{$\vec{u}_t$?}
%
%\begin{proof}
%Let $\mat{K}$ be a feedback controller such that $\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$. Let the output of
%the feedback system be a zero order holder such that $\vec{u}(t+kT_s)=\vec{u}(kT_s) : 0 \leq t < T_s$.
%The closed loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%\begin{equation}
%\vec{x}_{k,t}=\mat{A}_t\vec{x}_k-\mat{B}_t\mat{K}\vec{x}_k = (\mat{A}_t-\mat{B}_t\mat{K})\vec{x}_k : 0 \leq t < T_s
%\end{equation}
%We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
%\begin{align}
%\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
%\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
%\end{align}
%which accelerating results becomes
%\begin{align}
%\label{eq:feedback_sampled_cont}
%\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
%\vec{x}_{t} &= (\mat{A}_{t-kT_s}-\mat{B}_{t-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
%\label{eq:feedback_cont}
%\end{align}
%\end{proof}

@@

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
|{\vec{y}_e}_{k}|_1 \leq q_t, \text{ where } q_t=&(|\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})|_1+|\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}|_1)|U|_\infty
\end{align}
\end{theorem}
%
\todo{${\vec{y}_e}_{k}$ ?, notation $| |_x$ ?}
\todo{also, now we do have an output $y$, why didn't we consider this case at
\ref{sec:real_discrete_feedback_inputs}?}
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by the numerical representation} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop system meets given safety
specifications.  A safety specification gives raise to a requirement on the
states of the model, such that the feedback controller (namely the choice of
the gains matrix~$K$) must ensure that the state never violates the
requirement.  Note that a stable, closed-loop system is not necessarily a
safe system: indeed, the state values may leave the safe part of the state
space while they converge to the equilibrium, which is typical in the case
of oscillatory dynamics.  In~this work, the safety property is expressed as:
%
\todo{shouldn't this be over the states of the system?}
\begin{align}
\phi_\mathit{safety}\equiv& \underline{y} \leq y_k \leq \overline{y}
\iff \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
\label{eq:safetyspec}
\end{align}

where $\mat{G}$ is a matrix of template directions taken as row vectors \todo{unclear to me}, and $\vec{h}$ is an upper
bound for the set in the corresponding row direction.

@@

t_s<-log_{|\overline{\lambda}|}({p_s}) : |\overline{\lambda}| =\max(|\lambda_i|) \wedge |\overline{\lambda}|<1,\ \forall i \in [1\ n]

@@

%$t=-\frac{ln(d)}{|\overline{\omega}|}=-\frac{ln(d)}{ln(|\overline{\lambda}|)}$. 

@@

Careful examination of the above matrices reveals that because the $\mat{A}'$ is
upper block diagonal, the spectrum of $\mat{A}'$ is equal to the union of
the spectrum of $\mat{A}_{c}$ and $\mat{A}_{o}$, which can be calculated
separately.  Additionally, these correspond to the roots of the polynomials
%
\begin{align*}
\left(z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}\right) \left(z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}\right)
\end{align*}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Safety of a Digitally Observed System}\label{sec:cof_safety}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\todo{FWL effects here?}

>>>>

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable:
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^p, u \in \mathbb{R}^q, A \in \mathbb{R}^{p \times p}, B \in \mathbb{R}^{p \times q}, 
\end{align}
%
where $t \in \mathbb R_0^+$, where $A$ and $B$ are matrices that 
specify the continuous plant, and with initial states set as $x(0)$.

@@

In this work, equations~\eqref{eq:ode1} and~\eqref{eq:ode2} is soundly discretized in time (see Appendix~\ref{sec:cont_aa}) \todo{or use another paper as reference and remove sec .1?} into
\begin{align}
\label{eq:plant}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k
\end{align} 
%
where $k \in \mathbb N$ and $x_{0}=x(0)$ is the initial state. 
$\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ denote the matrices that describe the discretized plant dynamics, whereas $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ denote the continuous plant dynamics.  In this work we will assume $\mat{D}_d=\mat{D}=\mat{0}$. Since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of the model starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs restricted to $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big( \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big)\;. 
\end{equation}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe the safe state-space.
%
We extend the notation above to convex sets of initial conditions and inputs ($X_0$ and $U$), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x}_0 \in
X_0 \wedge \vec{u} \in U \wedge \vec{x} = \tau(\vec{x}_0,\vec{u})\}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore write $\tau^n(X_0,U)$ to denote the set of states reached from
$X_0$ via $\tau$ in $n$ steps (\emph{$n$-reach set}), i.e.\ for $n\geq 1$
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid &\vec{x}_0 \in X_0 \wedge \forall k\in [0,n-1]\nonumber\\ 
 &\vec{u}_{k} \in U \wedge \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and
vector sums preserve convexity, the sets $X_n = \tau^n(X_0,U)$ are also
convex.

@@

 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, we have
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \text{where }\mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{and } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  We call $\vec{x}_T$ a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 %% We can apply Th.~\ref{eq:continuous_tube_dyn} to discretized systems without inputs in order to
 %% obtain a witness ..
%
\todo{Can the corallary be motivated? The general motivation of the section was observing the system at times that are not necessarily integer multiple of the sampling time...} 
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0 : k \in \mathbb{R}^+$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in observing systems with parametric
 inputs, described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$.
 For such systems, a witness $\vec{x}_T$ is computed next.
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
\end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in \mathbb{R}^+$ is a witness of the system at time $kT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

@@

\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn}, 
 $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ and 
 $\mat{A}_d, \mat{B}_d$ describing the discretized plant dynamics is a witness of the system
 \begin{align}
 \dot{\vec{x}}(t) &= \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\nonumber\\
 \vec{u}(t)&=\mat{K}\vec{x}_k  \mid  kT_s \leq t < (k+1)T_s
 \end{align}
 \end{theorem}
%
%
%\begin{proof}
%Let $\mat{K}$ be a feedback controller such that $\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$. Let the output of
%the feedback system be a zero order holder such that $\vec{u}(t+kT_s)=\vec{u}(kT_s) : 0 \leq t < T_s$.
%The closed loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%\begin{equation}
%\vec{x}_{k,t}=\mat{A}_t\vec{x}_k-\mat{B}_t\mat{K}\vec{x}_k = (\mat{A}_t-\mat{B}_t\mat{K})\vec{x}_k : 0 \leq t < T_s
%\end{equation}
%We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
%\begin{align}
%\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
%\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
%\end{align}
%which accelerating results becomes
%\begin{align}
%\label{eq:feedback_sampled_cont}
%\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
%\vec{x}_{t} &= (\mat{A}_{t-kT_s}-\mat{B}_{t-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
%\label{eq:feedback_cont}
%\end{align}
%\end{proof}

@@

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
|\vec{y}_{k}-{\vec{y}_{t_d}}_{k}|_1 \leq q_t, \text{ where } q_t=&(|\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})|_1+|\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}|_1)|U|_\infty
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$.
\end{theorem}
%
\todo{notation $| |_x$ ?}
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by the numerical representation} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop system meets given safety
specifications.  A safety specification gives raise to a requirement on the
states of the model, such that the feedback controller (namely the choice of
the gains matrix~$K$) must ensure that the state never violates the
requirement.  Note that a stable, closed-loop system is not necessarily a
safe system: indeed, the state values may leave the safe part of the state
space while they converge to the equilibrium, which is typical in the case
of oscillatory dynamics.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}\equiv& \underline{y} \leq y_k \leq \overline{y}
\iff \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
\label{eq:safetyspec}
\end{align}

where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe the safe state-space.

@@

t_s<-log_{\overline{\lambda}}({p_s}) : \overline{\lambda} = \max(|\lambda_i|) \wedge \overline{\lambda}<1,\ \forall i \in [1\ n]

@@

%$t=-\frac{ln(d)}{|\overline{\omega}|}=-\frac{ln(d)}{ln(\overline{\lambda})}$. 

@@

Careful examination of the above matrices reveals that because the $\mat{A}'$ is
upper block diagonal, the spectrum of $\mat{A}'$ is equal to the union of
the spectrum of $\mat{A}_{c}$ and $\mat{A}_{o}$, which can be calculated
separately.  In other words, the characteristic polynomial of the closed loop is
equal to the multiplication of the characteristic polynomials of the controller
and the observer when seen independently.
%
\begin{align*}
P(z)=P_K(z)P_L(z)=\left(z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}\right) \left(z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}\right)
\end{align*}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Safety of a Digitally Observed System}\label{sec:cof_safety}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\todo{FWL effects here? they are implicit from equation \ref{eq:observer_LTI_FWL_cf}}
